\section{Blocks}
\label{sec:blocks}

A block is a header plus an ordered transaction vector. The header commits to
the previous block header, transaction Merkle root, timestamp, target encoding,
and nonce. The transaction vector begins with exactly one coinbase transaction.

\subsection{Block header}

A serialized \code{BlockHeader} is exactly 80 bytes in the field order shown:

\begin{longtable}{@{}L{0.18\linewidth}L{0.18\linewidth}L{0.22\linewidth}L{0.32\linewidth}@{}}
\toprule
\multicolumn{4}{@{}l}{\textbf{\code{BlockHeader}}} \\
\midrule
Field & Type & Encoding & Meaning \\
\midrule
\code{version} & \code{int32} & 4 little-endian bytes & Block version. Version bits use the high-bit pattern defined by \bip{9}. \\
{\small\code{previous}\newline\code{BlockHash}} & \code{uint256} & 32 bytes & Double-SHA256 hash of the previous block header in protocol byte order. \\
\code{merkleRoot} & \code{uint256} & 32 bytes & Merkle root of transaction identifiers. Witness data is committed separately. \\
\code{time} & \code{uint32} & 4 little-endian bytes & Miner-supplied Unix timestamp. \\
\code{bits} & \code{uint32} & 4 little-endian bytes & Compact proof-of-work target encoding. \\
\code{nonce} & \code{uint32} & 4 little-endian bytes & Nonce varied during proof-of-work search. \\
\bottomrule
\end{longtable}

For a block \(B\), \(\mathrm{bh}(B)=\mathrm{bh}(B.\text{\code{header}})\).

\subsection{Block record}

\begin{longtable}{@{}L{0.18\linewidth}L{0.20\linewidth}L{0.22\linewidth}L{0.30\linewidth}@{}}
\toprule
\multicolumn{4}{@{}l}{\textbf{\code{Block}}} \\
\midrule
Field & Type & Encoding & Meaning \\
\midrule
\code{header} & \code{BlockHeader} & 80 bytes & Serialized block header. \\
\code{txCount} & \code{compactSize} & CompactSize & Number of serialized transactions. \\
\code{transactions} & {\small\code{vector<}\newline\code{Transaction>}} & transaction encodings in order & First transaction must be coinbase; all later transactions must not be coinbase. \\
\bottomrule
\end{longtable}

When witness data is present, block serialization includes each transaction's
witness serialization for relay and block-weight computation. The transaction
identifier Merkle root continues to use \code{txid}, not \code{wtxid}.

\subsection{Merkle root}

For a block transaction vector \([tx_0,\ldots,tx_{n-1}]\), define:

\[
  L = [txid(tx_0), txid(tx_1), \ldots, txid(tx_{n-1})].
\]

A block commits to \(\mathrm{MR}(L)\). A block with empty \(L\) is invalid.

During root computation, a node rejects the duplicated-subtree mutation case:
if two identical hashes are paired at a level before an odd-last duplication
step, the block is invalid as mutated
\cite{bitcoin-core-v31}.

\subsection{Coinbase transaction}

The first transaction in each block is the coinbase transaction. It is the only
transaction permitted to have a null prevout. No other transaction in a block
may be coinbase. The coinbase creates the block subsidy and collects transaction
fees; its total output value must not exceed subsidy plus fees when the block is
connected.

\(\operatorname{Rules}(N,C,h,t)\) determines whether the candidate height must
be serialized at the beginning of the coinbase \code{scriptSig}
\cite{bip-0034,bitcoin-core-v31}.

\subsection{Witness commitment}

If \(R=\operatorname{Rules}(N,C,h,t)\) includes SegWit and any transaction has
witness data, the coinbase must commit to the witness Merkle root. The witness
tree leaves are:

\[
  W = [0^{256}, wtxid(tx_1), \ldots, wtxid(tx_{n-1})].
\]

The coinbase input witness contains a 32-byte reserved value \(r\). The
witness commitment is:

\[
  wc = \mathrm{H}(\mathrm{MR}(W) || r).
\]

The commitment appears in a coinbase output whose script begins with the
SegWit commitment header. If multiple outputs match, the highest-index matching
output is used \cite{bip-0141}.

\subsection{Context-free block checks}

\begin{longtable}{@{}L{0.28\linewidth}L{0.60\linewidth}@{}}
\toprule
Check & Rule \\
\midrule
Transaction count & The block contains at least one transaction. \\
Transaction count weight & \code{txCount * WITNESS\_SCALE\_FACTOR <= MAX\_BLOCK\_WEIGHT}. \\
Coinbase position & Transaction zero is coinbase; no later transaction is coinbase. \\
Merkle root & Header \code{merkleRoot} equals \(\mathrm{MR}(L)\) and is not mutated. \\
Stripped size & \code{strippedSize(block) * WITNESS\_SCALE\_FACTOR <= MAX\_BLOCK\_WEIGHT}. \\
Transaction validity & Every transaction passes context-free transaction checks. \\
Legacy sigops & \code{legacySigOps(block) * WITNESS\_SCALE\_FACTOR <= MAX\_BLOCK\_SIGOPS\_COST}. \\
\bottomrule
\end{longtable}

\subsection{Contextual block checks}

For a candidate block at height \(h\) extending chain \(C\), the previous block
is \(C[h-1]\), and \(R=\operatorname{Rules}(N,C,h,\code{time})\).

\begin{longtable}{@{}L{0.28\linewidth}L{0.60\linewidth}@{}}
\toprule
Check & Rule \\
\midrule
Previous header & The previous header exists and is valid for extension. \\
Median time & \code{time} is greater than \(\operatorname{MTP}(C,h-1)\). \\
Future time & A node does not accept the block while \code{time} is more than \code{MAX\_FUTURE\_BLOCK\_TIME} after the node's current clock time; this is temporary future status, not permanent consensus invalidity. \\
BIP94 timewarp & On networks enforcing BIP94, the first block of each difficulty-adjustment interval except genesis must have \code{time >= previous.time - BIP94\_TIMEWARP\_GRACE}. \\
Difficulty & \code{bits} encodes the target required for the height and network. \\
Finality & All transactions satisfy \(\operatorname{AbsFinal}(tx,h,t)\), where \(t=\operatorname{MTP}(C,h-1)\) if \(R\) includes BIP113 and the candidate block time otherwise. \\
Coinbase height & If \(R\) includes BIP34, the block commits to the correct height in the coinbase. \\
Witness commitment & If \(R\) includes SegWit and the block has witness data, the block contains the correct coinbase witness commitment. \\
Weight & Block weight is at most \code{MAX\_BLOCK\_WEIGHT}. \\
\bottomrule
\end{longtable}

The timestamp checks above follow the active validation predicates for median
time past, future-time acceptance, and BIP94 timewarp mitigation
\cite{bitcoin-core-v31,bip-0094}.
