\section{Ledger Model}
\label{sec:ledger-model}

Bitcoin validation is modeled as deterministic state transition over a
proof-of-work header graph, available block data, and the active UTXO set. This
section fixes the shared relations used by later sections.

\subsection{State}

A full-validation node's protocol state is:

\[
  S = (C,H,B,U,Q,\Pi)
\]

\begin{longtable}{@{}L{0.14\linewidth}L{0.24\linewidth}L{0.50\linewidth}@{}}
\toprule
Symbol & Type & Meaning \\
\midrule
\(C\) & \code{ActiveChain} & Ordered sequence of connected blocks from genesis to the active tip. \\
\(H\) & \code{HeaderIndex} & Known block-header graph keyed by block hash, including height, predecessor, target, timestamp, and accumulated work. \\
\(B\) & \code{BlockDataSet} & Full block data known to the node, keyed by block hash. \\
\(U\) & \code{UTXOSet} & Map from unspent outpoints to coin records at the active-chain tip. \\
\(Q\) & \code{Mempool} & Locally accepted unconfirmed transactions. This is relay state, not consensus state. \\
\(\Pi\) & \code{PeerSet} & Connected-peer state and negotiated network capabilities. \\
\bottomrule
\end{longtable}

Consensus validation reads \(C,H,B,U\) and mutates only \(C,H,B,U\). Mempool
and peer transitions read consensus state but do not define block validity.

\subsection{Coins}

The UTXO set maps outpoints to coins:

\[
  U[(txid,n)] = (value, scriptPubKey, height, coinbase).
\]

\begin{longtable}{@{}L{0.22\linewidth}L{0.24\linewidth}L{0.42\linewidth}@{}}
\toprule
Coin field & Type & Meaning \\
\midrule
\code{value} & \code{Amount} & Output amount in satoshis. Consensus-valid values are in the money range. \\
\code{scriptPubKey} & \code{Script} & Locking script committed by the creating transaction output. \\
\code{height} & \code{uint32} & Active-chain height of the creating transaction. \\
\code{coinbase} & \code{bool} & True iff the creating transaction is the block coinbase. \\
\bottomrule
\end{longtable}

\subsection{Rule context}

Let \(N\) be the selected network parameter set. For candidate height \(h\),
candidate block time \(t\), and active chain \(C\), define:

\[
  R = \operatorname{Rules}(N,C,h,t).
\]

\(R\) is the set of consensus rules active for the candidate object. It includes
network constants, buried activation heights, versionbits state, timestamp
rules, and script flags. Later sections use \(R\) rather than restating whether
P2SH, BIP34, strict DER, locktime, sequence locks, SegWit, Taproot, Tapscript,
or network-specific timewarp mitigations are active.

\subsection{Validity classes}

\begin{longtable}{@{}L{0.24\linewidth}L{0.64\linewidth}@{}}
\toprule
Class & Meaning \\
\midrule
\(\operatorname{DecodeValid}_{T}(b)\) & \(\mathrm{E}_{T}^{-1}(b)\) succeeds and consumes a canonical encoding of type \(T\). \\
\(\operatorname{FormValid}(x)\) & Object \(x\) is internally well-formed independent of chain state. \\
\(\operatorname{ContextValid}_{R}(x,C)\) & Object \(x\) satisfies height, time, and active-rule predicates under \(R\). \\
\(\operatorname{StateValid}_{R}(x,U)\) & Object \(x\) is valid against the current spendable coin set. \\
\(\operatorname{RelayValid}_{P}(x,Q,U)\) & Object \(x\) is accepted by local relay policy. This is not a consensus predicate. \\
\bottomrule
\end{longtable}

\subsection{Transition relations}

For a non-coinbase transaction, the transaction transition is:

\[
  \operatorname{ApplyTx}(R,U,tx) =
  \begin{cases}
    U', & \text{if } tx \text{ is valid against } U \text{ under } R,\\
    \bot, & \text{otherwise.}
  \end{cases}
\]

For a block \(B_h\) at height \(h\), block connection is:

\[
  \operatorname{ConnectBlock}(R,(C,U),B_h) =
  \begin{cases}
    (C',U'), & \text{if every block and transaction predicate succeeds,}\\
    \bot, & \text{otherwise.}
  \end{cases}
\]

Header acceptance, block-data acceptance, chain selection, and reorganization
are the consensus transitions defined in Section~\ref{sec:consensus}. Mempool
admission is the local transition defined in Section~\ref{sec:mempool}.

\subsection{Invariants}

Every accepted active-chain transition preserves:

\begin{longtable}{@{}L{0.24\linewidth}L{0.64\linewidth}@{}}
\toprule
Invariant & Statement \\
\midrule
Determinism & For fixed \(N,C,U,B_h\), successful connection produces a unique \(C',U'\). \\
Outpoint uniqueness & At any active tip, each outpoint appears at most once in \(U\). \\
No double spend & A connected transaction consumes each input outpoint at most once, and a connected block consumes each live outpoint at most once. \\
Conservation & A non-coinbase transaction cannot create value: \(\sum outputs(tx) \le \sum spent(tx)\). \\
Issuance bound & For each connected block, \(\sum outputs(coinbase) \le subsidy(h)+\sum fees(B_h)\). \\
Work selection & The active chain is selected from valid candidates by greatest accumulated work. \\
\bottomrule
\end{longtable}
