\appendixsection{Mainnet Parameters}
\label{app:parameters}

This appendix collects network parameters and historical consensus exceptions
that are inputs to \(\operatorname{Rules}(N,C,h,t)\). Mainnet is the default
network for this specification.

\subsection{Mainnet parameters}

\begin{longtable}{@{}L{0.30\linewidth}L{0.58\linewidth}@{}}
\toprule
Parameter & Value \\
\midrule
Genesis hash & {\scriptsize\code{000000000019d6689c085ae165831e93}\newline\code{4ff763ae46a2a6c172b3f1b60a8ce26f}} \\
Genesis Merkle root & {\scriptsize\code{4a5e1e4baab89f3a32518a88c31bc87f}\newline\code{618f76673e2cc77ab2127b7afdeda33b}} \\
Message start & \code{f9 be b4 d9} \\
Subsidy halving interval & \code{SUBSIDY\_HALVING\_INTERVAL} \\
Proof-of-work target spacing & \code{TARGET\_SPACING} \\
Proof-of-work target timespan & \code{TARGET\_TIMESPAN} \\
Bech32 human-readable part & \code{bc} \\
\bottomrule
\end{longtable}

Other networks, including testnet, testnet4, signet, and regtest, have distinct
genesis blocks, proof-of-work limits, retargeting exceptions, message-start
bytes, ports, address prefixes, and activation parameters. Their rules are not
inferred from mainnet.

\subsection{Historical exceptions}

\begin{longtable}{@{}L{0.28\linewidth}L{0.60\linewidth}@{}}
\toprule
Exception & Rule \\
\midrule
Genesis coinbase & The genesis block's coinbase transaction is part of the block history but its output is not inserted into \(U\) by normal block connection. \\
BIP30 duplicates & Historical duplicate-transaction cases are handled by the BIP30 no-overwrite rule and its deployed exceptions. Outside those exceptions, creating an output whose outpoint is already unspent is invalid. \\
Buried activations & Mainnet rules whose activation heights are buried in contemporary software are represented by \(\operatorname{Rules}(N,C,h,t)\), not by live versionbits state. \\
Future-dated blocks & A block whose timestamp is more than \code{MAX\_FUTURE\_BLOCK\_TIME} after the node's current clock is temporarily ineligible for active validation, not permanently invalid. \\
\bottomrule
\end{longtable}

Signet uses a block signature challenge in the witness commitment payload.
Testnet4 and regtest can enforce BIP94 timewarp mitigation through network
parameters; mainnet does not set that parameter in the Bitcoin Core 31.0
baseline \cite{bip-0030,bip-0034,bip-0094,bip-0325,bitcoin-core-v31}.
