\section{Script Execution}
\label{sec:script}

For each non-coinbase input \(i\), spend authorization is the predicate:

\[
  \operatorname{Authorize}_{R}(tx,i,coin,u,H) \in \{\top,\bot\},
\]

where \(u=(scriptSig,witness)\) is the unlocking data, \(coin\) is the spent
UTXO, \(R\) is the active rule context, and \(H\) is the candidate height and
median-time context. Authorization chooses a spend form, constructs a script
machine state, and evaluates the committed program.

\subsection{Spend forms}

\begin{longtable}{@{}L{0.24\linewidth}L{0.64\linewidth}@{}}
\toprule
Form & Required relation \\
\midrule
Legacy & \(\operatorname{Eval}_{R}(scriptSig,\mathrm{BASE})\) then \(\operatorname{Eval}_{R}(scriptPubKey,\mathrm{BASE})\); witness empty; final stack top is true. \\
P2SH & \(R\) includes P2SH; \code{scriptSig} is push-only; final pushed item is \code{redeemScript}; \(\operatorname{HASH160}(redeemScript)\) matches; \(\operatorname{Eval}_{R}(redeemScript,\mathrm{BASE})\) succeeds on the pushed stack. \\
Witness v0 key hash & Native witness program length 20; \code{scriptSig} empty unless P2SH-wrapped; witness stack is \((signature,pubkey)\); \(\operatorname{HASH160}(pubkey)\) matches; ECDSA verifies \(M_{\mathrm{wit0}}\). \\
Witness v0 script hash & Native witness program length 32; final witness item is \code{witnessScript}; \(\operatorname{SHA256}(witnessScript)\) matches; \(\operatorname{Eval}_{R}(witnessScript,\mathrm{WITNESS\_V0})\) succeeds on the remaining witness stack. \\
Taproot key path & \(R\) includes Taproot; witness stack, after optional annex removal, contains one Schnorr signature; BIP340 verifies \(M_{\mathrm{tr}}\) under the output key. \\
Taproot script path & Control block commits a tapleaf to the output key. Leaf version \code{0xc0} evaluates as Tapscript; unknown leaf versions succeed after commitment validation. \\
Unknown witness version & A syntactically valid witness program with no assigned meaning succeeds without execution. \\
\bottomrule
\end{longtable}

Native witness spends require empty \code{scriptSig}; P2SH-wrapped witness
spends require \code{scriptSig} to push exactly the witness program
\cite{bip-0016,bip-0141,bip-0143,bip-0340,bip-0341,bip-0342}.

\subsection{Machine state}

Script evaluation is a deterministic transition over:

\[
  X=(tx,i,coin,script,pc,S,A,C,R,\sigma,k,\alpha,\beta).
\]

\begin{longtable}{@{}L{0.18\linewidth}L{0.70\linewidth}@{}}
\toprule
Field & Meaning \\
\midrule
\code{script}, \code{pc} & Program bytes and instruction cursor. \\
\(S,A,C\) & Main stack, altstack, and conditional-execution stack. \\
\(R,\sigma\) & Active rule context and signature version: \(\mathrm{BASE}\), \(\mathrm{WITNESS\_V0}\), or \(\mathrm{TAPSCRIPT}\). \\
\(k\) & Last executed code-separator position for signature-message construction. \\
\(\alpha\) & Taproot annex, if present. \\
\(\beta\) & Tapscript validation-weight budget. \\
\bottomrule
\end{longtable}

Legacy and witness-v0 execution enforce the script limits named in
Appendix~\ref{app:constants}. Tapscript removes the script-size and
per-script-opcode limits, keeps the stack-item and stack-size limits, and sets:

\[
  \beta =
  \text{\code{TAPSCRIPT\_SIGOPS\_BUDGET\_BASE}}+
  \operatorname{serializedWitnessSize}(input).
\]

Each executed signature opcode with a nonempty signature subtracts
\code{TAPSCRIPT\_SIGOPS\_COST}; negative \(\beta\) fails.

\subsection{Evaluation}

\[
  \operatorname{Eval}_{R}(script,\sigma,S_0)=
  \begin{cases}
    \top, & \text{execution halts successfully with required final stack,}\\
    \bot, & \text{parsing, execution, resource, or final-stack validation fails.}
  \end{cases}
\]

If \(\sigma=\mathrm{TAPSCRIPT}\), the script is scanned for any
\code{OP\_SUCCESSx} index before ordinary execution; encountering one succeeds.
Otherwise, instructions are decoded left to right and \(\operatorname{Step}\)
is applied while all entries of \(C\) are true. Non-control opcodes in inactive
branches do not mutate \(S\), but parse-stage failures and Tapscript
\code{OP\_SUCCESSx} detection still apply.

Successful termination requires an empty \(C\). Legacy and witness-v0 require a
nonempty \(S\) whose top element casts to true. Tapscript requires exactly one
stack element, and that element must cast to true.

\subsection{Primitive relations}

\begin{longtable}{@{}L{0.28\linewidth}L{0.60\linewidth}@{}}
\toprule
Relation & Definition \\
\midrule
\(\operatorname{CastToBool}(x)\) & False iff \(x\) is empty, zero, or negative zero; true otherwise. \\
\(\operatorname{Num}(x,n,min)\) & Signed-magnitude little-endian integer from byte vector \(x\), length at most \(n\); if \(min\), nonzero encodings must be minimal. \\
\(\operatorname{MinimalPush}(op,x)\) & True iff \(op\) is the shortest permitted push form for \(x\). \\
\(\operatorname{LockTime}(X,n)\) & Units match \code{tx.lockTime}, \(n\le tx.lockTime\), and the input sequence is nonfinal. \\
\(\operatorname{Sequence}(X,n)\) & BIP68 is active in \(R\), disable/type bits permit comparison, and the input sequence satisfies \(n\). \\
\(\operatorname{CheckSig}(X,sig,key,h)\) & Empty signatures return false. Legacy and witness-v0 use ECDSA over secp256k1; Taproot and Tapscript use BIP340 Schnorr. Tapscript key size 0 fails, size 32 verifies, and other nonzero key sizes are unknown key types that pass current consensus. \\
\bottomrule
\end{longtable}

\subsection{Signature messages}

Signature checks verify exactly one digest:

\[
\begin{aligned}
  M_{\mathrm{legacy}} &= \operatorname{SigMsg}_{\mathrm{legacy}}(tx,i,k,h),\\
  M_{\mathrm{wit0}} &= \operatorname{SigMsg}_{\mathrm{wit0}}(tx,i,coin,k,h),\\
  M_{\mathrm{tr}} &= \operatorname{SigMsg}_{\mathrm{tr}}(tx,i,coin,\alpha,k,h).
\end{aligned}
\]

Hash type bytes are \code{0x00} \code{DEFAULT} for Taproot,
\code{0x01} \code{ALL}, \code{0x02} \code{NONE}, \code{0x03}
\code{SINGLE}, and \code{0x80} \code{ANYONECANPAY} ORed with non-default
forms. \code{ALL}, \code{NONE}, and \code{SINGLE} select the output commitment
set; \code{ANYONECANPAY} restricts input commitment to the signed input.
Legacy out-of-range \code{SINGLE} signs \code{uint256::ONE}; Taproot
out-of-range \code{SINGLE} fails \cite{bip-0143,bip-0341,bip-0342}.

\subsection{Opcode classes}

Effects use \code{(before -- after)} for stack transitions, with the rightmost
item at the top of \(S\). Parameterized direct byte pushes are represented as a
single family; every named opcode is listed separately.

\begin{longtable}{@{}L{0.14\linewidth}L{0.34\linewidth}L{0.40\linewidth}@{}}
\toprule
Index & Opcode & Effect \\
\midrule
\code{0} & \code{OP\_0 / OP\_FALSE} & Push empty vector. \\
\code{1-75} & \code{OP\_PUSHBYTES\_n} & Push the next \(n\) script bytes. \\
\code{76} & \code{OP\_PUSHDATA1} & Next \code{uint8} gives pushed byte length. \\
\code{77} & \code{OP\_PUSHDATA2} & Next little-endian \code{uint16} gives pushed byte length. \\
\code{78} & \code{OP\_PUSHDATA4} & Next little-endian \code{uint32} gives pushed byte length. \\
\code{79} & \code{OP\_1NEGATE} & Push \(-1\). \\
\code{81} & \code{OP\_1} & Push integer \(1\). \\
\code{82} & \code{OP\_2} & Push integer \(2\). \\
\code{83} & \code{OP\_3} & Push integer \(3\). \\
\code{84} & \code{OP\_4} & Push integer \(4\). \\
\code{85} & \code{OP\_5} & Push integer \(5\). \\
\code{86} & \code{OP\_6} & Push integer \(6\). \\
\code{87} & \code{OP\_7} & Push integer \(7\). \\
\code{88} & \code{OP\_8} & Push integer \(8\). \\
\code{89} & \code{OP\_9} & Push integer \(9\). \\
\code{90} & \code{OP\_10} & Push integer \(10\). \\
\code{91} & \code{OP\_11} & Push integer \(11\). \\
\code{92} & \code{OP\_12} & Push integer \(12\). \\
\code{93} & \code{OP\_13} & Push integer \(13\). \\
\code{94} & \code{OP\_14} & Push integer \(14\). \\
\code{95} & \code{OP\_15} & Push integer \(15\). \\
\code{96} & \code{OP\_16} & Push integer \(16\). \\
\bottomrule
\end{longtable}

\begin{longtable}{@{}L{0.14\linewidth}L{0.34\linewidth}L{0.40\linewidth}@{}}
\toprule
Index & Opcode & Effect \\
\midrule
\code{97} & \code{OP\_NOP} & No effect. \\
\code{99} & \code{OP\_IF} & \code{(v --)}; push \(\operatorname{CastToBool}(v)\) to \(C\). \\
\code{100} & \code{OP\_NOTIF} & \code{(v --)}; push negated condition to \(C\). \\
\code{101} & \code{OP\_VERIF} & Fail. \\
\code{102} & \code{OP\_VERNOTIF} & Fail. \\
\code{103} & \code{OP\_ELSE} & Toggle top of \(C\); unmatched \code{ELSE} fails. \\
\code{104} & \code{OP\_ENDIF} & Pop top of \(C\); unmatched \code{ENDIF} fails. \\
\code{105} & \code{OP\_VERIFY} & \code{(v --)} if true; otherwise fail. \\
\code{106} & \code{OP\_RETURN} & Fail. \\
\bottomrule
\end{longtable}

\begin{longtable}{@{}L{0.14\linewidth}L{0.34\linewidth}L{0.40\linewidth}@{}}
\toprule
Index & Opcode & Effect \\
\midrule
\code{107} & \code{OP\_TOALTSTACK} & Move top of \(S\) to \(A\). \\
\code{108} & \code{OP\_FROMALTSTACK} & Move top of \(A\) to \(S\). \\
\code{109} & \code{OP\_2DROP} & \code{(x1 x2 --)}. \\
\code{110} & \code{OP\_2DUP} & \code{(x1 x2 -- x1 x2 x1 x2)}. \\
\code{111} & \code{OP\_3DUP} & \code{(x1 x2 x3 -- x1 x2 x3 x1 x2 x3)}. \\
\code{112} & \code{OP\_2OVER} & \code{(x1 x2 x3 x4 -- x1 x2 x3 x4 x1 x2)}. \\
\code{113} & \code{OP\_2ROT} & \code{(x1 x2 x3 x4 x5 x6 -- x3 x4 x5 x6 x1 x2)}. \\
\code{114} & \code{OP\_2SWAP} & \code{(x1 x2 x3 x4 -- x3 x4 x1 x2)}. \\
\code{115} & \code{OP\_IFDUP} & Duplicate top item iff it casts true. \\
\code{116} & \code{OP\_DEPTH} & Push stack depth. \\
\code{117} & \code{OP\_DROP} & \code{(x --)}. \\
\code{118} & \code{OP\_DUP} & \code{(x -- x x)}. \\
\code{119} & \code{OP\_NIP} & \code{(x1 x2 -- x2)}. \\
\code{120} & \code{OP\_OVER} & \code{(x1 x2 -- x1 x2 x1)}. \\
\code{121} & \code{OP\_PICK} & \code{(n -- x)}; copy stack item \(n\) to the top. \\
\code{122} & \code{OP\_ROLL} & \code{(n -- x)}; move stack item \(n\) to the top. \\
\code{123} & \code{OP\_ROT} & \code{(x1 x2 x3 -- x2 x3 x1)}. \\
\code{124} & \code{OP\_SWAP} & \code{(x1 x2 -- x2 x1)}. \\
\code{125} & \code{OP\_TUCK} & \code{(x1 x2 -- x2 x1 x2)}. \\
\code{130} & \code{OP\_SIZE} & \code{(x -- x len(x))}. \\
\bottomrule
\end{longtable}

\begin{longtable}{@{}L{0.14\linewidth}L{0.34\linewidth}L{0.40\linewidth}@{}}
\toprule
Index & Opcode & Effect \\
\midrule
\code{135} & \code{OP\_EQUAL} & Byte-vector equality. \\
\code{136} & \code{OP\_EQUALVERIFY} & Equality followed by \code{VERIFY}. \\
\code{139} & \code{OP\_1ADD} & \code{(x -- x+1)} over \(\operatorname{Num}(x,4,min)\). \\
\code{140} & \code{OP\_1SUB} & \code{(x -- x-1)} over \(\operatorname{Num}(x,4,min)\). \\
\code{143} & \code{OP\_NEGATE} & \code{(x -- -x)} over \(\operatorname{Num}(x,4,min)\). \\
\code{144} & \code{OP\_ABS} & \code{(x -- |x|)} over \(\operatorname{Num}(x,4,min)\). \\
\code{145} & \code{OP\_NOT} & Push \(1\) iff numeric input is zero; otherwise push \(0\). \\
\code{146} & \code{OP\_0NOTEQUAL} & Push \(0\) iff numeric input is zero; otherwise push \(1\). \\
\code{147} & \code{OP\_ADD} & \code{(x y -- x+y)} over script numbers. \\
\code{148} & \code{OP\_SUB} & \code{(x y -- x-y)} over script numbers. \\
\code{154} & \code{OP\_BOOLAND} & Push \(1\) iff both numeric inputs are nonzero. \\
\code{155} & \code{OP\_BOOLOR} & Push \(1\) iff at least one numeric input is nonzero. \\
\code{156} & \code{OP\_NUMEQUAL} & Numeric equality. \\
\code{157} & \code{OP\_NUMEQUALVERIFY} & Numeric equality followed by \code{VERIFY}. \\
\code{158} & \code{OP\_NUMNOTEQUAL} & Numeric inequality. \\
\code{159} & \code{OP\_LESSTHAN} & Numeric less-than comparison. \\
\code{160} & \code{OP\_GREATERTHAN} & Numeric greater-than comparison. \\
\code{161} & \code{OP\_LESSTHANOREQUAL} & Numeric less-than-or-equal comparison. \\
\code{162} & \code{OP\_GREATERTHANOREQUAL} & Numeric greater-than-or-equal comparison. \\
\code{163} & \code{OP\_MIN} & Push the smaller numeric input. \\
\code{164} & \code{OP\_MAX} & Push the larger numeric input. \\
\code{165} & \code{OP\_WITHIN} & \code{(x min max -- ok)} where \(min \le x < max\). \\
\bottomrule
\end{longtable}

\begin{longtable}{@{}L{0.14\linewidth}L{0.34\linewidth}L{0.40\linewidth}@{}}
\toprule
Index & Opcode & Effect \\
\midrule
\code{166} & \code{OP\_RIPEMD160} & \code{(x -- RIPEMD160(x))}. \\
\code{167} & \code{OP\_SHA1} & \code{(x -- SHA1(x))}. \\
\code{168} & \code{OP\_SHA256} & \code{(x -- SHA256(x))}. \\
\code{169} & \code{OP\_HASH160} & \code{(x -- HASH160(x))}. \\
\code{170} & \code{OP\_HASH256} & \code{(x -- H(x))}. \\
\code{171} & \code{OP\_CODESEPARATOR} & Set \(k\) to this opcode position. \\
\code{172} & \code{OP\_CHECKSIG} & \code{(sig key -- ok)} using \(\operatorname{CheckSig}\). \\
\code{173} & \code{OP\_CHECKSIGVERIFY} & \code{CHECKSIG} followed by \code{VERIFY}. \\
\code{174} & \code{OP\_CHECKMULTISIG} & \code{(dummy sig[m] m key[n] n -- ok)}; enforces historical dummy rule when active; disabled in Tapscript. \\
\code{175} & \code{OP\_CHECKMULTISIGVERIFY} & \code{CHECKMULTISIG} followed by \code{VERIFY}; disabled in Tapscript. \\
\code{186} & \code{OP\_CHECKSIGADD} & Tapscript only: \code{(sig n key -- n+ok)}. \\
\bottomrule
\end{longtable}

\begin{longtable}{@{}L{0.14\linewidth}L{0.34\linewidth}L{0.40\linewidth}@{}}
\toprule
Index & Opcode & Effect \\
\midrule
\code{176} & \code{OP\_NOP1} & No effect. \\
\code{177} & \code{OP\_CHECKLOCKTIMEVERIFY / OP\_NOP2} & If active in \(R\), require \(\operatorname{LockTime}(X,n)\); otherwise no effect. \\
\code{178} & \code{OP\_CHECKSEQUENCEVERIFY / OP\_NOP3} & If active in \(R\), require \(\operatorname{Sequence}(X,n)\); otherwise no effect. \\
\code{179} & \code{OP\_NOP4} & No effect unless redefined by active soft-fork rules. \\
\code{180} & \code{OP\_NOP5} & No effect unless redefined by active soft-fork rules. \\
\code{181} & \code{OP\_NOP6} & No effect unless redefined by active soft-fork rules. \\
\code{182} & \code{OP\_NOP7} & No effect unless redefined by active soft-fork rules. \\
\code{183} & \code{OP\_NOP8} & No effect unless redefined by active soft-fork rules. \\
\code{184} & \code{OP\_NOP9} & No effect unless redefined by active soft-fork rules. \\
\code{185} & \code{OP\_NOP10} & No effect unless redefined by active soft-fork rules. \\
\bottomrule
\end{longtable}

\begin{longtable}{@{}L{0.14\linewidth}L{0.34\linewidth}L{0.40\linewidth}@{}}
\toprule
Index & Opcode & Effect \\
\midrule
\code{80} & \code{OP\_RESERVED / OP\_SUCCESS80} & Reserved failure if executed outside Tapscript; Tapscript success. \\
\code{98} & \code{OP\_VER / OP\_SUCCESS98} & Reserved failure if executed outside Tapscript; Tapscript success. \\
\code{126} & \code{OP\_CAT / OP\_SUCCESS126} & Disabled failure outside Tapscript; Tapscript success. \\
\code{127} & \code{OP\_SUBSTR / OP\_SUCCESS127} & Disabled failure outside Tapscript; Tapscript success. \\
\code{128} & \code{OP\_LEFT / OP\_SUCCESS128} & Disabled failure outside Tapscript; Tapscript success. \\
\code{129} & \code{OP\_RIGHT / OP\_SUCCESS129} & Disabled failure outside Tapscript; Tapscript success. \\
\code{131} & \code{OP\_INVERT / OP\_SUCCESS131} & Disabled failure outside Tapscript; Tapscript success. \\
\code{132} & \code{OP\_AND / OP\_SUCCESS132} & Disabled failure outside Tapscript; Tapscript success. \\
\code{133} & \code{OP\_OR / OP\_SUCCESS133} & Disabled failure outside Tapscript; Tapscript success. \\
\code{134} & \code{OP\_XOR / OP\_SUCCESS134} & Disabled failure outside Tapscript; Tapscript success. \\
\code{137} & \code{OP\_RESERVED1 / OP\_SUCCESS137} & Reserved failure if executed outside Tapscript; Tapscript success. \\
\code{138} & \code{OP\_RESERVED2 / OP\_SUCCESS138} & Reserved failure if executed outside Tapscript; Tapscript success. \\
\code{141} & \code{OP\_2MUL / OP\_SUCCESS141} & Disabled failure outside Tapscript; Tapscript success. \\
\code{142} & \code{OP\_2DIV / OP\_SUCCESS142} & Disabled failure outside Tapscript; Tapscript success. \\
\code{149} & \code{OP\_MUL / OP\_SUCCESS149} & Disabled failure outside Tapscript; Tapscript success. \\
\code{150} & \code{OP\_DIV / OP\_SUCCESS150} & Disabled failure outside Tapscript; Tapscript success. \\
\code{151} & \code{OP\_MOD / OP\_SUCCESS151} & Disabled failure outside Tapscript; Tapscript success. \\
\code{152} & \code{OP\_LSHIFT / OP\_SUCCESS152} & Disabled failure outside Tapscript; Tapscript success. \\
\code{153} & \code{OP\_RSHIFT / OP\_SUCCESS153} & Disabled failure outside Tapscript; Tapscript success. \\
\code{187} & \code{OP\_SUCCESS187} & Tapscript success; invalid outside Tapscript. \\
\code{188} & \code{OP\_SUCCESS188} & Tapscript success; invalid outside Tapscript. \\
\code{189} & \code{OP\_SUCCESS189} & Tapscript success; invalid outside Tapscript. \\
\code{190} & \code{OP\_SUCCESS190} & Tapscript success; invalid outside Tapscript. \\
\code{191} & \code{OP\_SUCCESS191} & Tapscript success; invalid outside Tapscript. \\
\code{192} & \code{OP\_SUCCESS192} & Tapscript success; invalid outside Tapscript. \\
\code{193} & \code{OP\_SUCCESS193} & Tapscript success; invalid outside Tapscript. \\
\code{194} & \code{OP\_SUCCESS194} & Tapscript success; invalid outside Tapscript. \\
\code{195} & \code{OP\_SUCCESS195} & Tapscript success; invalid outside Tapscript. \\
\code{196} & \code{OP\_SUCCESS196} & Tapscript success; invalid outside Tapscript. \\
\code{197} & \code{OP\_SUCCESS197} & Tapscript success; invalid outside Tapscript. \\
\code{198} & \code{OP\_SUCCESS198} & Tapscript success; invalid outside Tapscript. \\
\code{199} & \code{OP\_SUCCESS199} & Tapscript success; invalid outside Tapscript. \\
\code{200} & \code{OP\_SUCCESS200} & Tapscript success; invalid outside Tapscript. \\
\code{201} & \code{OP\_SUCCESS201} & Tapscript success; invalid outside Tapscript. \\
\code{202} & \code{OP\_SUCCESS202} & Tapscript success; invalid outside Tapscript. \\
\code{203} & \code{OP\_SUCCESS203} & Tapscript success; invalid outside Tapscript. \\
\code{204} & \code{OP\_SUCCESS204} & Tapscript success; invalid outside Tapscript. \\
\code{205} & \code{OP\_SUCCESS205} & Tapscript success; invalid outside Tapscript. \\
\code{206} & \code{OP\_SUCCESS206} & Tapscript success; invalid outside Tapscript. \\
\code{207} & \code{OP\_SUCCESS207} & Tapscript success; invalid outside Tapscript. \\
\code{208} & \code{OP\_SUCCESS208} & Tapscript success; invalid outside Tapscript. \\
\code{209} & \code{OP\_SUCCESS209} & Tapscript success; invalid outside Tapscript. \\
\code{210} & \code{OP\_SUCCESS210} & Tapscript success; invalid outside Tapscript. \\
\code{211} & \code{OP\_SUCCESS211} & Tapscript success; invalid outside Tapscript. \\
\code{212} & \code{OP\_SUCCESS212} & Tapscript success; invalid outside Tapscript. \\
\code{213} & \code{OP\_SUCCESS213} & Tapscript success; invalid outside Tapscript. \\
\code{214} & \code{OP\_SUCCESS214} & Tapscript success; invalid outside Tapscript. \\
\code{215} & \code{OP\_SUCCESS215} & Tapscript success; invalid outside Tapscript. \\
\code{216} & \code{OP\_SUCCESS216} & Tapscript success; invalid outside Tapscript. \\
\code{217} & \code{OP\_SUCCESS217} & Tapscript success; invalid outside Tapscript. \\
\code{218} & \code{OP\_SUCCESS218} & Tapscript success; invalid outside Tapscript. \\
\code{219} & \code{OP\_SUCCESS219} & Tapscript success; invalid outside Tapscript. \\
\code{220} & \code{OP\_SUCCESS220} & Tapscript success; invalid outside Tapscript. \\
\code{221} & \code{OP\_SUCCESS221} & Tapscript success; invalid outside Tapscript. \\
\code{222} & \code{OP\_SUCCESS222} & Tapscript success; invalid outside Tapscript. \\
\code{223} & \code{OP\_SUCCESS223} & Tapscript success; invalid outside Tapscript. \\
\code{224} & \code{OP\_SUCCESS224} & Tapscript success; invalid outside Tapscript. \\
\code{225} & \code{OP\_SUCCESS225} & Tapscript success; invalid outside Tapscript. \\
\code{226} & \code{OP\_SUCCESS226} & Tapscript success; invalid outside Tapscript. \\
\code{227} & \code{OP\_SUCCESS227} & Tapscript success; invalid outside Tapscript. \\
\code{228} & \code{OP\_SUCCESS228} & Tapscript success; invalid outside Tapscript. \\
\code{229} & \code{OP\_SUCCESS229} & Tapscript success; invalid outside Tapscript. \\
\code{230} & \code{OP\_SUCCESS230} & Tapscript success; invalid outside Tapscript. \\
\code{231} & \code{OP\_SUCCESS231} & Tapscript success; invalid outside Tapscript. \\
\code{232} & \code{OP\_SUCCESS232} & Tapscript success; invalid outside Tapscript. \\
\code{233} & \code{OP\_SUCCESS233} & Tapscript success; invalid outside Tapscript. \\
\code{234} & \code{OP\_SUCCESS234} & Tapscript success; invalid outside Tapscript. \\
\code{235} & \code{OP\_SUCCESS235} & Tapscript success; invalid outside Tapscript. \\
\code{236} & \code{OP\_SUCCESS236} & Tapscript success; invalid outside Tapscript. \\
\code{237} & \code{OP\_SUCCESS237} & Tapscript success; invalid outside Tapscript. \\
\code{238} & \code{OP\_SUCCESS238} & Tapscript success; invalid outside Tapscript. \\
\code{239} & \code{OP\_SUCCESS239} & Tapscript success; invalid outside Tapscript. \\
\code{240} & \code{OP\_SUCCESS240} & Tapscript success; invalid outside Tapscript. \\
\code{241} & \code{OP\_SUCCESS241} & Tapscript success; invalid outside Tapscript. \\
\code{242} & \code{OP\_SUCCESS242} & Tapscript success; invalid outside Tapscript. \\
\code{243} & \code{OP\_SUCCESS243} & Tapscript success; invalid outside Tapscript. \\
\code{244} & \code{OP\_SUCCESS244} & Tapscript success; invalid outside Tapscript. \\
\code{245} & \code{OP\_SUCCESS245} & Tapscript success; invalid outside Tapscript. \\
\code{246} & \code{OP\_SUCCESS246} & Tapscript success; invalid outside Tapscript. \\
\code{247} & \code{OP\_SUCCESS247} & Tapscript success; invalid outside Tapscript. \\
\code{248} & \code{OP\_SUCCESS248} & Tapscript success; invalid outside Tapscript. \\
\code{249} & \code{OP\_SUCCESS249} & Tapscript success; invalid outside Tapscript. \\
\code{250} & \code{OP\_SUCCESS250} & Tapscript success; invalid outside Tapscript. \\
\code{251} & \code{OP\_SUCCESS251} & Tapscript success; invalid outside Tapscript. \\
\code{252} & \code{OP\_SUCCESS252} & Tapscript success; invalid outside Tapscript. \\
\code{253} & \code{OP\_SUCCESS253} & Tapscript success; invalid outside Tapscript. \\
\code{254} & \code{OP\_SUCCESS254} & Tapscript success; invalid outside Tapscript. \\
\code{255} & \code{OP\_INVALIDOPCODE} & Fail. \\
\bottomrule
\end{longtable}

Consensus script behavior is defined by the active P2SH, DER, locktime,
sequence, NULLDUMMY, SegWit, Taproot, and Tapscript rules
\cite{bip-0065,bip-0066,bip-0068,bip-0112,bip-0113,bip-0147,bip-0341,bip-0342,bitcoin-core-v31}.
